Forecasting future oil production in Norway and the UK: a general improved methodology

### 2014-07-10 Publications/Forecasting  Future Oil Production/peak_oil_6.tex
Latest version.

%% LyX 2.0.2 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[twoside]{article}
\usepackage{mathpazo}
\renewcommand{\ttdefault}{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{float}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{esint}
\usepackage{url}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 1},backref=section,colorlinks=false]
 {hyperref}
\usepackage{breakurl}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

\@ifundefined{date}{}{\date{}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.2 (15/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

@@

\usepackage{fancyhdr}% Headers and footers
 % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyhead[C]{Forecasting future oil production $\bullet$ July 2014} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

@@

@@

\author{L. Fi\'evet, Z. Forr\'o, %M. del Degan, 
P. Cauwels and D. Sornette\\
ETH Z\"urich\\
Department of Management, Technology and Economics\\
Chair of Entrepreneurial Risks}

%\date{}
\maketitle
\begin{abstract}
\noindent We present a new Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. based on a two-step process,
(i) the nonlinear extrapolation of the current/past
performances of individual oil fields and (ii) a stochastic model
of the frequency of future oil field discoveries. Compared
with the standard methodology that tends to underestimate remaining oil reserves,
our method gives a better description of future oil production, as validated by
our back-tests starting in 2008. Specifically, we predict remaining reserves
extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $90 \pm10$ million
barrels for the UK, which are respectively $45\%$ and $66\%$ above the
predictions using the standard methodology.
\end{abstract}

@@

\lettrine[nindent=0em,lines=3] Forecasting future oil production
has been a topic of active interest since the beginning of the past
century because of oil central role in our economy. Its importance
ranges from energy production, manufacturing to the pharmaceutical
industry. As petroleum is a non-renewable and finite resource, it
is primordial to be able to forecast future oil production. The fear
of a global oil peak, beyond which production will inevitably decline,
has been growing due to stagnating supplies and high oil prices since
the crisis in 2008/2009 \cite{Murray 2012}. As any industrialized
country, Europe is strongly dependent on oil supply to maintain its
economic power. In the nowadays difficult geopolitical environment,
it is important to know how much of the oil needed in Europe will
come from reliable sources. In the past, a big share has been coming
from Norway and the U.K., two of Europe's biggest exporters. However,
the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
production has been declining rapidly as well \cite{H=0000F6=0000F6k 2008}.\\
 The methodology behind forecasting future oil production has not
evolved much since M. King Hubbert, who in 1956 famously predicted
that the U.S. oil production would peak around 1965-1970 \cite{Hubbert 1956}.
That prediction has proven itself to be correct. His main argument
was based on the finiteness of oil reserves and to what amounts
as the use of the logistic differential equation for the total
quantity $P(t)$ of oil extracted up to time $t$
\begin{equation}
\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)~.
\label{logistic}
\end{equation}
The logistic differential equation
is characterized by an initial exponential growth, which then decreases
to zero as the total oil extracted reaches saturation (no more oil
is to be found). The parameter $r$ is commonly referred to as the
growth rate, and $K$ as the carrying capacity (total quantity of
oil that can be ultimately extracted). If $P(t)$ is the amount of oil extracted up to
time $t$, then $f(t) := \frac{dP}{dt}$ is the oil production rate, the quantity
that M. King Hubbert predicted with surprising accuracy to peak. From
a methodological point of view, the Hubbert model has enjoyed a longstanding
popularity in modeling future oil production given its simplicity.
Various extensions have been studied by Brandt \cite{Brandt 2006}
to account for multi-cycled or asymmetric production curves.
The existing forecasts of future oil production use some form of the
Hubbert model (Brecha, Laherr\`{e}re or Lynch \cite{Brecha 2012,Laherr=0000E8re 2002,Lynch 2002})
or some economical model applied to aggregate production (Greiner \cite{Greiner 2012}),
but none goes into the details of studying the underlying dynamics.
The main reason for the lack of details is certainly the lack of available
data.

In this article, a new methodology is introduced to forecast future
oil production. Instead of taking the aggregate oil production profile
and fitting it with the Hubbert curve or its variants (such as the
multi-cyclic Hubbert curve), the production profile of each individual
oil field is used. By extending their production into the future and
extrapolating the future rate of discovery of new fields, the future
oil production is forecasted by means of a Monte Carlo simulation.
To demonstrate the generality of the methodology presented here,
it is applied to two major oil producing countries with publicly available
data: Norway\cite{Norwegian Data} and the U.K. \cite{U.K. Data}.

@@

To be able to forecast the oil production of each individual field,
regularity had to be found in the production's dynamics. Modeling
the whole production profile from the beginning of extraction seems
elusive due to the variety of the forms it can take. Fortunately,
modeling the decay process is sufficient in order to extrapolate future
oil production. A preliminary classification is necessary to achieve
that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
show that, independent of the country, oil fields can be classified
into three main categories:
\begin{itemize}
\item \textbf{Regular fields} - Their decays show some regularity (see figure \ref{regular} ); 
\item \textbf{Irregular fields} - The ones that do not decay in a regular
fashion (see figure \ref{irregular});
\item \textbf{New fields} - The ones that do not decay yet. As such, there
is no easy way to forecast their future oil production based on past
data (see figure \ref{new}).
\end{itemize}
\noindent All the fields have been fitted using an automated algorithm,
but the results have been subsequently checked visually to sort out
the irregular fields which could not be fitted. As of January 2014,
regular fields make up 85\% and 87\% of the number of fields and 94\%
and 71\% of the total produced oil volume in Norway and the U.K. respectively.
As such, being able to model them is crucial. To capture as many different
decay dynamics as possible, the decay part of the oil production rate
$f(t) := \frac{dP}{dt}$ has been fitted by the stretched exponential  
\begin{equation}
f(t)=f_{0}~e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
\label{stretched_exponential}
\end{equation}
The stretched exponential function
has many advantages as it generalises the power law and can therefore
capture a broad variety of distributions as shown by Laherr\`{e}re and
Sornette \cite{Laherr=0000E8re and Sornette 1998}. Moreover,
Malevergne et al. \cite{malpissor05} showed that the power law
function can be obtained as an asymptotic case of the stretched
exponential family, allowing for asymptotically nested statistical tests.  As can be seen
in figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
is a good functional form to fit the decay process of regular fields.

@@

A natural choice for such a model is a non-homogenous Poisson process.
The Poisson process is a process that generates independent events
at a rate $\lambda$. It is nonhomogeneous if the rate is time-dependent,
$\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
is to find a functional form for $N(t)$, the statistical average
of the cumulative number of events (discoveries) up to time $t$.
Then, $\lambda(t)$ is simply a smoothed estimation 
of the observed rate $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
shows $N(t)$ for Norwegian fields classified according to their size in two classes, 
dwarfs and giant fields. 
The logistic curve is a good fit to the data (integral form of equation
\ref{logistic}). This implies that after an initial increase, the
rate of new discoveries reaches a peak followed by a decrease until
no more oil fields are to be found, consistent with our fundamental
observations. This same approach has already been successfully applied
by Forr\'o et al. \cite{Forro 2012} to estimate the number of daily active
users on Zynga.

As the discovery and production dynamics are not independent of the
field size, the fields have been split into two groups: dwarfs and
giants. Unfortunately, the two logistic curves thus obtained are highly
sensitive to the splitting size. This results from the major issue,
when fitting a logistic curve to data, that the carrying capacity
can not be determined if the data does not already exhibit the slowdown 
in growth towards the carrying capacity. However, it is
mentioned in the literature that often dwarf fields have already been
discovered a long time ago, but their production has been postponed
for economical reasons \cite[p. 378]{Lynch 2002}. Therefore, it is
expected that the large oil fields have mostly been found and produced,
and that future discoveries will mostly be made up of dwarf fields.
Consequently, the splitting size has been picked as small as possible
in order to maximize the number of giant fields but large enough to
avoid recent discoveries. Our definition is thus:
\begin{itemize}
\item \textbf{Dwarfs}: Fields which produced less than $50\cdot10^{6}$
barrels.
\item \textbf{Giants}: Fields which produced more than $50\cdot10^{6}$
barrels.
\end{itemize}
We note that this definition differs by a factor 10 from the more standard one, for which
oil fields with an ultimate recoverable resource of 0.5 billion barrels (Gb) 
or higher are classified as giants, while oil fields with smaller URR are considered to be dwarfs
\cite{H=0000F6=0000F6k 2008}.

The resulting plot shown in figure \ref{rate_of_discovery} pictures
the dynamics: giant oil fields have mostly been found while the discovery
process for dwarf fields is still ongoing. The logistic growth curve
fit to the giant discoveries is well constrained, however the fit
for the dwarfs is poorly constrained. As can be seen in figure \ref{rate_of_discovery},
the carrying capacity $K$ of the logistic growth model is not well
constrained by the available data. A large spectrum of values for
$K$ can lead to an equally good fit of the data. We have not taken
into account the possible effect that
newly discovered dwarf fields could become smaller and smaller 
until the new fields have a size that is so small to be economical to drill.

There are in fact two competing effects that are likely to compensate each other.
On the one hand, figure \ref{ccdffieldsizes} shows the complementary 
cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
Two salient properties can be observed. First, the tails 
of the distributions are well described by power laws
\begin{equation}
{\rm CCDF}(S) \sim {1 \over S^{1+\alpha}}~,
\label{ccdfsizefield}
\end{equation}
with exponent $\alpha = 1.2 \pm 0.1$ for Norway and $\alpha = 1.4 \pm 0.1$ for the UK.
The fact that the estimations of the exponents $\alpha$ are larger than $1$ implies
that the cumulative oil reserves are asymptotically controlled by the largest fields, and not the small ones
\cite{Sornettecritbook}. However, the fact that the exponents $\alpha$ are rather close
to $1$ (which is called ``Zipf law'') would make the many small oil field contributing 
significantly in total. This brings us to the second important feature exhibited by figure \ref{ccdffieldsizes},
namely the roll-overs of the CCDFs for small oil fields, likely due to an under-sampling
of the data. Indeed, as for most data sets involving broad distributions of sizes such
as oil fields, the distributions are in general incomplete for the small events due to the non
exhaustive sampling. This incompleteness raises the spectre that our extrapolations
might be grossly underestimating the large potential contributions to the total reserves
of the many small yet undiscovered small oil fields. Assuming that the power law (\ref{ccdfsizefield})
would hold for smaller fields down to size of 1 Million barrels leads to a number of 
such fields 10 to 100 times larger than presently known.

But there is another
key factor that needs to be considered, namely the fact that small oil fields
are not economically viable for exploitation, which leads to an effective truncation
in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
Taking the data from the NPD\cite{Norwegian Data}
providing the yearly investments broken by fields in Norway, let us consider the illustrative
case of the field GAUPE. The investment spent to 
develop its exploitation was  \$380M, while its estimated size is 
$\simeq 1.2Mb$, which corresponds approximately to a total
market value of $\$120M$ at \$100 per barrel. It is thus not a surprise that
investment to exploit this field was interrupted in 2013. 
Figure \ref{returnNorwayfield} shows the total estimated revenue divided
by investment as a function of the estimated ultimate recovery for a number
of small oil fields. It can be inferred that oil fields of sizes smaller than about 10 Mb
are not economically viable as long as the market oil price does not 
grow much higher than \$100 per barrel. This implies that, notwithstanding
the large number of unknown small oil fields, economic considerations
oblige us to neglect the small oil fields, therefore providing
a justification of our procedure. In fact, economic constraints may lead to 
cap the carrying capacity at a value smaller than 
the one shown in figure  \ref{rate_of_discovery}.

@@

To overcome the poor constraint on the carrying capacity
$K$ obtained from the fitting procedure for dwarf fields, a method
already used by Smith \cite{Smith 1980} has been implemented. This
method makes the following two postulates:
\begin{enumerate}
\item ``The discovery of reservoirs in a petroleum play can be modeled
statistically as sampling without replacement from the underlying
population of reservoirs.''
\item ``The discovery of a particular reservoir from among the existing
population is random, with a probability of discovery being dependent 
on (proportional to) reservoir size.''
\end{enumerate}
The fields are split into $J$ size bins denoted $S_{1},\,\ldots,\, S_{J}$
occurring with frequency $n_{1},\,\ldots,\, n_{J}$. Each discovery
is considered as a step $i$ at which a field of size $I(i)\in\{S_{1},\,\ldots,\, S_{J}\}$
is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
discovered before the $i^{th}$ step. Then, the probability that the
discovery at step $i$ is of size $j$ can be expressed as
\begin{equation}
P\left(I(i)=S_{j}\right)=\frac{\left(n_{j}-m_{ij}\right)\cdot S_{j}}{\sum_{k=1}^{J}\left(n_{k}-m_{ik}\right)\cdot S_{k}}.
\end{equation}
The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
can then be expressed as
\begin{equation}
L=\prod_{i=1}^{N}\frac{\left(n_{I(i)}-m_{iI(i)}\right)\cdot S_{I(i)}}{\sum_{j=1}^{J}\left(n_{j}-m_{ij}\right)\cdot S_{j}}.\label{eq:likelihood}
\end{equation}
The unknown parameters are the number of fields $n_{1},\,\ldots,\, n_{J}$,
whose likelihood can now be estimated based on the existing discoveries.
Using a brute force approach, the entire space of plausible values
for the variables $n_{1},\,\ldots,\, n_{J}$ has been sampled. The
values $n_{j}$ have been sampled between the number of existing fields
$m_{Nj}$ in the bin $j$ and up to a value $n_{j}^{upper}$, such
that the scenario with the largest likelihood according to equation
(\ref{eq:likelihood}) has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
the bin $j$. Subsequently, the likelihood of each scenario (value
of the tuple $n_{1},\,\ldots,\, n_{J}$) has been normalized such
that the total likelihood of all generated scenarios equals one.

@@

Superposing the production curves results in the expected production
curve from future oil field for a given scenario.

@@

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels 
predicted by the extrapolation of the  \textbf{Fit}
of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Model}}$.}
\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $31\%$\tabularnewline
U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $40\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

The results in table \ref{forecast-14} show a striking difference
between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $30\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue for us to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$40\%$ compared to the latter.

@@

@@

%FIGURE 5
\begin{figure}[p]
\includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
\caption{Logistic fit of the function solution of expression (\ref{logistic})
to the number of discoveries for Norway. The discovery
rate of new oil fields is dependent on their size as explained in the main text.}
\label{rate_of_discovery} 
\end{figure}

>>>>

\documentclass[review]{elsarticle}
\usepackage{mathpazo}
\renewcommand{\ttdefault}{mathpazo}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{fancyhdr}
\pagestyle{fancy}
\usepackage{float}
\usepackage{amstext}
\usepackage{graphicx}
\usepackage{esint}
\usepackage{url}
\usepackage[unicode=true,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 1},backref=false,colorlinks=false]
 {hyperref}
\usepackage{breakurl}

@@

\providecommand{\tabularnewline}{\\}

\@ifundefined{date}{}{\date{}}

@@

 % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyhead[C]{Forecasting future oil production $\bullet$ July 2014} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

@@

\defcitealias{Commission2014}{European Commission}
\defcitealias{Directorate2014}{NPD}

@@

\author[eth-mtec]{L. Fi\'evet\corref{cor1}}
\ead{lfievet@ethz.ch}

\author[eth-mtec]{Z. Forr\'o}
\ead{zforro@ethz.ch}

\author[eth-mtec]{P. Cauwels}
\ead{pcauwels@ethz.ch}

\author[eth-mtec]{D. Sornette}
\ead{dsornette@ethz.ch}

\cortext[cor1]{Corresponding author. Tel.: +41 44 632 83 79.}

\address[eth-mtec]{Chair of Entrepreneurial Risks, ETH Z\"urich, Scheuchzerstrasse 7 (SEC F), CH-8092 Z\"urich, Switzerland}

\maketitle
\begin{abstract}
\noindent We present a new Monte-Carlo methodology to forecast
the crude oil production of Norway and the U.K. based on a two-step process,
(i) the nonlinear extrapolation of the current/past
performances of individual oil fields and (ii) a stochastic model
of the frequency of future oil field discoveries. Compared
with the standard methodology that tends to underestimate remaining oil reserves,
our method gives a better description of future oil production, as validated by
our back-tests starting in 2008. Specifically, we predict remaining reserves
extractable until 2030 to be $188 \pm 10$ million barrels for Norway and $98 \pm10$ million
barrels for the UK, which are respectively $45\%$ and $66\%$ above the
predictions using the standard methodology.
\end{abstract}

@@

\lettrine[nindent=0em,lines=3] Forecasting future oil production
has been a topic of active interest since the beginning of the past
century because of oil central role in our economy. Its importance
ranges from energy production, manufacturing to the pharmaceutical
industry. As petroleum is a non-renewable and finite resource, it
is primordial to be able to forecast future oil production. The fear
of a global oil peak, beyond which production will inevitably decline,
has been growing due to stagnating supplies and high oil prices since
the crisis in 2008/2009 \citep{Murray2013}. As any industrialized
country, Europe is strongly dependent on oil supply to maintain its
economic power. In the nowadays difficult geopolitical environment,
it is important to know how much of the oil needed in Europe will
come from reliable sources. In the past, a big share has been coming
from Norway and the U.K., two of Europe's biggest exporters. However,
the U.K. already became a net importer in 2005 and Norway\textquoteright{}s
production has been declining rapidly as well \citep{Hook2008}.\\
 The methodology behind forecasting future oil production has not
evolved much since M. King Hubbert, who in 1956 famously predicted
that the U.S. oil production would peak around 1965-1970 \citep{Hubbert1956}.
That prediction has proven itself to be correct. His main argument
was based on the finiteness of oil reserves and to what amounts
as the use of the logistic differential equation for the total
quantity $P(t)$ of oil extracted up to time $t$
\begin{equation}
\frac{dP}{dt}=rP\left(1-\frac{P}{K}\right)~.
\label{logistic}
\end{equation}
The logistic differential equation
is characterized by an initial exponential growth, which then decreases
to zero as the total oil extracted reaches saturation (no more oil
is to be found). The parameter $r$ is commonly referred to as the
growth rate, and $K$ as the carrying capacity (total quantity of
oil that can be ultimately extracted). If $P(t)$ is the amount of oil extracted up to
time $t$, then $f(t) := \frac{dP}{dt}$ is the oil production rate, the quantity
that M. King Hubbert predicted with surprising accuracy to peak. From
a methodological point of view, the Hubbert model has enjoyed a longstanding
popularity in modeling future oil production given its simplicity.
Various extensions have been studied by \citet{Brandt2007}
to account for multi-cycled or asymmetric production curves.
The existing forecasts of future oil production use some form of the
Hubbert model \citep{Brecha2012,Laherrere2002,Lynch2002}
or some economical model applied to aggregate production \citep{Greiner2011},
but none goes into the details of studying the underlying dynamics.
The main reason for the lack of details is certainly the lack of available
data.

In this article, a new methodology is introduced to forecast future
oil production. Instead of taking the aggregate oil production profile
and fitting it with the Hubbert curve or its variants (such as the
multi-cyclic Hubbert curve), the production profile of each individual
oil field is used. By extending their production into the future and
extrapolating the future rate of discovery of new fields, the future
oil production is forecasted by means of a Monte Carlo simulation.
To demonstrate the generality of the methodology presented here,
it is applied to two major oil producing countries with publicly available
data: Norway \citepalias{Directorate2014} and the U.K. \citep{GOVUK2014}.

@@

To be able to forecast the oil production of each individual field,
regularity had to be found in the production's dynamics. Modeling
the whole production profile from the beginning of extraction seems
elusive due to the variety of the forms it can take. Fortunately,
modeling the decay process is sufficient in order to extrapolate future
oil production. A preliminary classification is necessary to achieve
that goal. Figures \ref{regular}, \ref{irregular} and \ref{new}
show that, independent of the country, oil fields can be classified
into three main categories:
\begin{itemize}
\item \textbf{Regular fields} - Their decays show some regularity (see figure \ref{regular} ); 
\item \textbf{Irregular fields} - The ones that do not decay in a regular
fashion (see figure \ref{irregular});
\item \textbf{New fields} - The ones that do not decay yet. As such, there
is no easy way to forecast their future oil production based on past
data (see figure \ref{new}).
\end{itemize}
\noindent All the fields have been fitted using an automated algorithm,
but the results have been subsequently checked visually to sort out
the irregular fields which could not be fitted. As of January 2014,
regular fields make up 85\% and 87\% of the number of fields and 94\%
and 71\% of the total produced oil volume in Norway and the U.K. respectively.
As such, being able to model them is crucial. To capture as many different
decay dynamics as possible, the decay part of the oil production rate
$f(t) := \frac{dP}{dt}$ has been fitted by the stretched exponential  
\begin{equation}
f(t)=f_{0}~e^{-\left(\frac{t}{\tau}\right)^{\beta}}~.
\label{stretched_exponential}
\end{equation}
The stretched exponential function
has many advantages as it generalises the power law and can therefore
capture a broad variety of distributions as shown by \citet{Laherrere1998}. Moreover,
\citet{Malevergne2005} showed that the power law
function can be obtained as an asymptotic case of the stretched
exponential family, allowing for asymptotically nested statistical tests.  As can be seen
in figure \ref{regular}, the stretched exponential (equation \ref{stretched_exponential})
is a good functional form to fit the decay process of regular fields.

@@

A natural choice for such a model is a non-homogenous Poisson process.
The Poisson process is a process that generates independent events
at a rate $\lambda$. It is nonhomogeneous if the rate is time-dependent,
$\lambda\rightarrow\lambda(t)$. The standard way to measure $\lambda(t)$
is to find a functional form for $N(t)$, the statistical average
of the cumulative number of events (discoveries) up to time $t$.
Then, $\lambda(t)$ is simply a smoothed estimation 
of the observed rate $\frac{dN(t)}{dt}$. Figure \ref{rate_of_discovery}
shows $N(t)$ for Norwegian fields classified according to their size in two classes, 
dwarfs and giant fields. 
The logistic curve is a good fit to the data (integral form of equation
\ref{logistic}). This implies that after an initial increase, the
rate of new discoveries reaches a peak followed by a decrease until
no more oil fields are to be found, consistent with our fundamental
observations. This same approach has already been successfully applied
by \cite{Forro2012} to estimate the number of daily active
users on Zynga.

As the discovery and production dynamics are not independent of the
field size, the fields have been split into two groups: dwarfs and
giants. Unfortunately, the two logistic curves thus obtained are highly
sensitive to the splitting size. This results from the major issue,
when fitting a logistic curve to data, that the carrying capacity
can not be determined if the data does not already exhibit the slowdown 
in growth towards the carrying capacity. However, it is
mentioned in the literature that often dwarf fields have already been
discovered a long time ago, but their production has been postponed
for economical reasons \cite[p. 378]{Lynch2002}. Therefore, it is
expected that the large oil fields have mostly been found and produced,
and that future discoveries will mostly be made up of dwarf fields.
Consequently, the splitting size has been picked as small as possible
in order to maximize the number of giant fields but large enough to
avoid recent discoveries. Our definition is thus:
\begin{itemize}
\item \textbf{Dwarfs}: Fields which produced less than $50\cdot10^{6}$
barrels.
\item \textbf{Giants}: Fields which produced more than $50\cdot10^{6}$
barrels.
\end{itemize}
We note that this definition differs by a factor 10 from the more standard one, for which
oil fields with an ultimate recoverable resource of 0.5 billion barrels (Gb) 
or higher are classified as giants, while oil fields with smaller URR are considered to be dwarfs
\citep{Hook2008}.

The resulting plot shown in figure \ref{rate_of_discovery} pictures
the dynamics: giant oil fields have mostly been found while the discovery
process for dwarf fields is still ongoing. The logistic growth curve
fit to the giant discoveries is well constrained, however the fit
for the dwarfs is poorly constrained. As can be seen in figure \ref{rate_of_discovery},
the carrying capacity $K$ of the logistic growth model is not well
constrained by the available data. A large spectrum of values for
$K$ can lead to an equally good fit of the data. We have not taken
into account the possible effect that
newly discovered dwarf fields could become smaller and smaller 
until the new fields have a size that is too small to be economical to drill.

There are in fact two competing effects that are likely to compensate each other.
On the one hand, figure \ref{ccdffieldsizes} shows the complementary 
cumulative distribution function (CCDF) of known oil field sizes $S$ from Norway and the UK.
Two salient properties can be observed. First, the tails 
of the distributions are well described by power laws
\begin{equation}
{\rm CCDF}(S) \sim {1 \over S^{1+\alpha}}~,
\label{ccdfsizefield}
\end{equation}
with exponent $\alpha = 1.2 \pm 0.1$ for Norway and $\alpha = 1.4 \pm 0.1$ for the UK.
The fact that the estimations of the exponents $\alpha$ are larger than $1$ implies
that the cumulative oil reserves are asymptotically controlled by the largest fields, and not the small ones
\citep{Sornette2004}. However, the fact that the exponents $\alpha$ are rather close
to $1$ (which is called ``Zipf law'') would make the many small oil field contributing 
significantly in total. This brings us to the second important feature exhibited by figure \ref{ccdffieldsizes},
namely the roll-overs of the CCDFs for small oil fields, likely due to an under-sampling
of the data. Indeed, as for most data sets involving broad distributions of sizes such
as oil fields, the distributions are in general incomplete for the small events due to the non
exhaustive sampling. This incompleteness raises the spectre that our extrapolations
might be grossly underestimating the large potential contributions to the total reserves
of the many small yet undiscovered small oil fields. Assuming that the power law (\ref{ccdfsizefield})
would hold for smaller fields down to size of 1 Million barrels leads to a number of 
such fields 10 to 100 times larger than presently known.

But there is another
key factor that needs to be considered, namely the fact that small oil fields
are not economically viable for exploitation, which leads to an effective truncation
in the distribution (\ref{ccdfsizefield}) relevant for the estimation of recoverable oil.
Taking the data from the \citetalias{Directorate2014}
providing the yearly investments broken by fields in Norway, let us consider the illustrative
case of the field GAUPE. The investment spent to 
develop its exploitation was  \$380M, while its estimated size is 
$\simeq 1.2Mb$, which corresponds approximately to a total
market value of $\$120M$ at \$100 per barrel. It is thus not a surprise that
investment to exploit this field was interrupted in 2013. 
Figure \ref{returnNorwayfield} shows the total estimated revenue divided
by investment as a function of the estimated ultimate recovery for a number
of small oil fields. It can be inferred that oil fields of sizes smaller than about 10 Mb
are not economically viable as long as the market oil price does not 
grow much higher than \$100 per barrel. This implies that, notwithstanding
the large number of unknown small oil fields, economic considerations
oblige us to neglect the small oil fields, therefore providing
a justification of our procedure. In fact, economic constraints may lead to 
cap the carrying capacity at a value smaller than 
the one shown in figure  \ref{rate_of_discovery}.

@@

To overcome the poor constraint on the carrying capacity
$K$ obtained from the fitting procedure for dwarf fields, a method
already used by \cite{Smith1980} has been implemented. This
method makes the following two postulates:
\begin{enumerate}
\item ``The discovery of reservoirs in a petroleum play can be modeled
statistically as sampling without replacement from the underlying
population of reservoirs.''
\item ``The discovery of a particular reservoir from among the existing
population is random, with a probability of discovery being dependent 
on (proportional to) reservoir size.''
\end{enumerate}
The fields are split into $J$ size bins denoted $S_{1},\,\ldots,\, S_{J}$
occurring with frequency $n_{1},\,\ldots,\, n_{J}$. Each discovery
is considered as a step $i$ at which a field of size $I(i)\in\{S_{1},\,\ldots,\, S_{J}\}$
is found and $m_{ij}$ denotes the number of fields of size $j\in\{1,\,\ldots,\, J\}$
discovered before the $i^{th}$ step. Then, the probability that the
discovery at step $i$ is of size $j$ can be expressed as
\begin{equation}
P\left(I(i)=S_{j}\right)=\frac{\left(n_{j}-m_{ij}\right)\cdot S_{j}}{\sum_{k=1}^{J}\left(n_{k}-m_{ik}\right)\cdot S_{k}}.
\end{equation}
The likelihood $L$ for a complete sequence of $N$ discoveries $\{I(1),\,\ldots,\, I(N)\}$
can then be expressed as
\begin{equation}
L=\prod_{i=1}^{N}\frac{\left(n_{I(i)}-m_{iI(i)}\right)\cdot S_{I(i)}}{\sum_{j=1}^{J}\left(n_{j}-m_{ij}\right)\cdot S_{j}}.\label{eq:likelihood}
\end{equation}
The unknown parameters are the number of fields $n_{1},\,\ldots,\, n_{J}$,
whose likelihood can now be estimated based on the existing discoveries.
Using a brute force approach, the entire space of plausible values
for the variables $n_{1},\,\ldots,\, n_{J}$ has been sampled. The
values $n_{j}$ have been sampled between the number of existing fields
$m_{Nj}$ in the bin $j$ and up to a value $n_{j}^{upper}$, such
that the scenario with the largest likelihood according to equation
(\ref{eq:likelihood}) has $n_{j}^{max}=n_{j}^{upper}/2$ fields in
the bin $j$. Subsequently, the likelihood of each scenario (value
of the tuple $n_{1},\,\ldots,\, n_{J}$) has been normalized such
that the total likelihood of all generated scenarios equals one.

@@

Superposing the production curves results in the expected production
curve from future oil fields for a given scenario.

@@

\begin{table}[H]
\caption{Remaining oil reserves until 2030 in barrels 
predicted by the extrapolation of the  \textbf{Fit}
of the past country production, and predicted by the Monte-Carlo \textbf{Model}.
The relative difference between these two predictions 
is defined by $\Delta=\frac{\textrm{Model}-\textrm{Fit}}{\textrm{Fit}}$.}
\centering %
\begin{tabular}{lllr}
 & Fit (barrels) & Model (barrels)  & $\Delta$ \tabularnewline
\midrule Norway & $130\cdot10^{6}$  & $188\cdot10^{6}$  & $45\%$\tabularnewline
U.K. & $59\cdot10^{6}$  & $98\cdot10^{6}$  & $66\%$ \tabularnewline
\bottomrule  &  &  & \tabularnewline
\end{tabular}\label{forecast-14}
\end{table}

The results in table \ref{forecast-14} show a striking difference
between the extrapolation of the fit and the Monte-Carlo model forecast. According to the
fit (extrapolation of aggregate production), Norway's future oil production
would decay much faster than in the Monte-Carlo case. The remaining
reserves estimated with the Monte-Carlo methodology are $45\%$ larger
then the estimate from the fit. This difference originates from two
different effects:
\begin{itemize}
\item The sum of the forecast of the individual existing fields is larger
than the extrapolation of the aggregate production.
\item The extrapolation of the aggregate production does not capture well
the discovery process of dwarf fields.
\end{itemize}
In the U.K., oil production faced a change of regime during the early
nineties due to technological innovation, giving rise to the inverted
\textquotedbl{}w shape\textquotedbl{} of the oil production profile.
However, this has not been an issue for us to extrapolate the decay of production
starting at the second peak. The difference between the Hubbert-based
methodology and the Monte-Carlo one is very similar to the Norwegian
case. The former underestimates the remaining oil reserves by about
$66\%$ compared to the latter.

@@

\pagebreak

@@

%FIGURE 5
\begin{figure}[H]
\includegraphics[width=1\columnwidth]{extrapolating-no-discoveries}
\caption{Logistic fit of the function solution of expression (\ref{logistic})
to the number of discoveries for Norway. The discovery
rate of new oil fields is dependent on their size as explained in the main text.}
\label{rate_of_discovery} 
\end{figure}

\pagebreak
